\vspace{-1.3cm} \textmd{\normalsize{High Performance Computing Project}}\\ Parallelization of the Probabilistic Roadmap Method with GPU Acceleration

### 2016-03-30 paper/paper.tex
paper fast fertig

\subsection{Project Overview}

@@

Algorithms 1 and 2 two show in pseudo code, how the work is split between host and device.

\begin{algorithm}
	\KwData{qs, qe, N, offset}
	\KwResult{res}
	compute distances $d_j = \abs{q_{end}^j - q_{start}^j}$ \;
	testnum[j] = $d_j/{\Delta q}$: number of threads for each line \;
	testpos[j] = thread displacements \;
	reset res \;
	call kernel \;
	\caption{Host indicator function}
\end{algorithm}

\begin{algorithm}
	\KwData{arrays qs, qe, testpos, testnum and threadindex}
	\KwResult{res}
	compute associated edge number $j$, defined by $testpos[j] \leq threadindex \le testpos[j]+testnum[j]$ \;
	$c = (threadindex - testpos[j]) / (testnum[k]-1) $ \;
	$q=c q_{start}^j + (1-c) q_{end}^j $ \;
	calculate robot kinematics $T_0(q), ..., T_d(q) $\;
	$res_{tmp} = 0; $ \;
	\For{all registered pairs pairs of polytopes $P, Q$}{
		$res_{tmp}=seperating\_vector\_algorithm(P, T_{i_P}, Q, T_{i_Q})$ \;
		\If{$res_{tmp} \neq 0$}{
			$res[j]=res_{tmp}$ \;
			break \;
		}
	}
	\caption{Robot kernel: code for one thread}
\end{algorithm}

@@

\section{Structure of the PRM Solver}\label{solver}

@@

\begin{lstlisting}
struct graph{
    std::map<int,block*> map; //map for accessing blocks
    std::vector<block> blocks;

    std::vector<float> qstorage; //length ndof*N

    int newblockpos;  //position of next new block
    int blocknum;     //number of used blocks
    .
    .
    .
};
\end{lstlisting}

@@

Tests have been made for a scenario with a robot with 4 joints. For creating and plotting the environment and the resulting trajectories, Matlab functions are provided.

- timings of different versions at same conditions

- difficult to test because

	- not deterministic

	- dependant on density of nodes: how many neighbors have to be tested per connection

\section{Conclusion}

>>>>

\newpage
\subsection{Project Overview}

@@

Algorithms \ref{host} and \ref{kernel} two show in pseudo code, how the work is split between host and device.

\begin{algorithm}\label{host}
	\KwData{qs, qe, N, offset}
	\KwResult{res}
	compute distances $d_j = \abs{q_{end}^j - q_{start}^j}$ \;
	testnum[j] = $d_j/{\Delta q}$: number of threads for each line \;
	testpos[j] = thread displacements \;
	reset res \;
	call kernel \;
	\caption{Host indicator function}
\end{algorithm}

\begin{algorithm}\label{kernel}
	\KwData{arrays qs, qe, testpos, testnum and threadindex}
	\KwResult{res}
	compute associated edge number $j$, defined by $testpos[j] \leq threadindex \le testpos[j]+testnum[j]$ \;
	$c = (threadindex - testpos[j]) / (testnum[k]-1) $ \;
	$q=c q_{start}^j + (1-c) q_{end}^j $ \;
	calculate robot kinematics $T_0(q), ..., T_d(q) $\;
	$res_{tmp} = 0; $ \;
	\For{all registered pairs pairs of polytopes $P, Q$}{
		$res_{tmp}=seperating\_vector\_algorithm(P, T_{i_P}, Q, T_{i_Q})$ \;
		\If{$res_{tmp} \neq 0$}{
			$res[j]=res_{tmp}$ \;
			break \;
		}
	}
	\caption{Robot kernel: code for one thread}
\end{algorithm}

@@

\newpage
\section{The PRM Solver}\label{solver}

@@

\begin{lstlisting}
struct graph{
    std::map<int,block*> map; //map for accessing blocks
    std::vector<block> blocks;
    std::vector<float> qstorage; //length ndof*N
    int newblockpos;  //position of next new block
    int blocknum;     //number of used blocks
    .
    .
};
\end{lstlisting}

@@

Tests have been made for a scenario with a robot with 4 joints. For creating and plotting the environment and the resulting trajectories, Matlab functions are provided.
Because of the dependancy on random numbers, the algorithm takes different amounts of time for different runs.
Therefore, 10 runs were made for each version and number of processes.
However, the different versions lead to a different average amount of GPU threads needed.
So, to be able to better compare how they scale, the time per number of GPU threads used as a quantity.
Table \ref{resulttable} shows the results.
One can see, that while the versions with node communication has a better time to GPU threads ratio, but leads to a greater amount of threads on average.
The versions with less communication however have better absolute times.
Furthermore, one sees the improvement from version 2 to version 3 from the asynchronous indicator function.

\ \\ 
\begin{table}[h]\label{resulttable}
\centering
\begin{tabular}{| c | c | r | r | r |}

\hline
version	&  processes/	& time (ms)	& GPU  		& time per 1e6  \\ 
		& GPUs		& 			& threads		& GPU threads (ms) \\ 
\hline
		& 1			& 1002		& 696413		& 1448	\\ 
1		& 2			& 763		& 838938		& 918	\\ 
		& 4			& 706		& 1113781	& 647	\\ 
\hline
		& 1			& 900		& 608172		& 1490	\\ 
2		& 2			& 647		& 608172		& 1070	\\ 
		& 4			& 568		& 608172		& 947	\\ 
\hline
		& 1			& 696		& 599847		& 1181	\\ 		
3		& 2			& 554		& 599847		& 951	\\ 
		& 4			& 500		& 599847		& 864	\\ 
\hline

\end{tabular}
\caption{Timing results for the parallelization versions from chapter \ref{solver}}
\end{table}
